---
title: "Hands-on Exercise 04"
author: "Leow Xian Zu"
date: "15 Sep 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true 
  message: false #no more warning message
  freeze: true #whatever document already commited does not render
---

# Set up

## Load packages

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, knitr)
```

## Importing data

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

## Left join

```{r}
hunan <- left_join(hunan, hunan2012) %>%
  dplyr::select(1:4, 7, 15)
```

# Visualisation of development indicator

```{r}
basemap <- tm_shape(hunan) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan, "GDPPC")
tmap_arrange(basemap, gdppc, asp=1, ncol=2)
```

## **Computing QUEEN**

```{r}
wm_q <- poly2nb(hunan, queen=TRUE)
summary(wm_q)
```

The summary report above shows that there are 88 area units in Hunan. The most connected area unit has 11 neighbours. There are two area units with only one heighbours.

For each polygon in our polygon object, *wm_q* lists all neighboring polygons. For example, to see the neighbors for the first polygon in the object, type:

```{r}
wm_q[[1]]
```

Polygon 1 has 5 neighbors. The numbers represent the polygon IDs as stored in hunan SpatialPolygonsDataFrame class.

We can retrive the county name of Polygon ID=1 by using the code chunk below:

```{r}
hunan$County[1]

```

To reveal the county names of the five neighboring polygons, the code chunk will be used, then we can retrieve the GDPPC of these five countries by using the code chunk below.

```{r}
hunan$NAME_3[c(2,3,4,57,85)]
nb1 <- wm_q[[1]]
nb1 <- hunan$GDPPC[nb1]
nb1
```

## Computing ROOK

```{r}
wm_r <- poly2nb(hunan, queen=FALSE)
summary(wm_r)
```

### **Visualising contiguity weights**

```{r}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
coords <- cbind(longitude, latitude)
head(coords)
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red")
```

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

## A look at queen and rook contiguity

```{r}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey", main="Queen Contiguity")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red")
plot(hunan$geometry, border="lightgrey", main="Rook Contiguity")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

# Computing distance based neighbours

## Cut off distance

```{r}
#coords <- coordinates(hunan)
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```

## Computing fixed distance weight matrix

```{r}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62

```

The value `3.681818` means that, on average, each region in the dataset has approximately 3.68 neighbours. This is derived by dividing the total number of links (324) by the number of regions (88). Some regions may have more or fewer neighbours, but on average, they each have about 3.68 connections with other regions.

```{r}
table(hunan$County, card(wm_d62))
```

#### Plotting fixed distance weight matrix

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE)
plot(k1, coords, add=TRUE, col="red", length=0.08)
```

```{r}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey", main="1st nearest neighbours")
plot(k1, coords, add=TRUE, col="red", length=0.08)
plot(hunan$geometry, border="lightgrey", main="Distance link")
plot(wm_d62, coords, add=TRUE, pch = 19, cex = 0.6)
```

### **Computing adaptive distance weight matrix**

```{r}
knn6 <- knn2nb(knearneigh(coords, k=6))
knn6
plot(hunan$geometry, border="lightgrey")
plot(knn6, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

## Weights based on IDW

```{r}
dist <- nbdists(wm_q, coords, longlat = TRUE)
ids <- lapply(dist, function(x) 1/(x))
ids
```

## **Row-standardised Weights Matrix**

```{r}
rswm_q <- nb2listw(wm_q, style="W", zero.policy = TRUE)
rswm_q
```

```{r}
rswm_q$weights[10]
```

Each neighbor is assigned a 0.125 of the total weight. This means that when R computes the average neighboring income values, each neighbor's income will be multiplied by 0.125 before being tallied.

Using the same method, we can also derive a row standardised distance weight matrix by using the code chunk below.

```{r}
rswm_ids <- nb2listw(wm_q, glist=ids, style="B", zero.policy=TRUE)
rswm_ids

```

```{r}
rswm_ids$weights[1]
summary(unlist(rswm_ids$weights))
```

## **Application of Spatial Weight Matrix**

### **Spatial lag with row-standardized weights**

```{r}
GDPPC.lag <- lag.listw(rswm_q, hunan$GDPPC)
GDPPC.lag
```

```{r}
nb1 <- wm_q[[1]]
nb1 <- hunan$GDPPC[nb1]
nb1
```

```{r}
lag.list <- list(hunan$NAME_3, lag.listw(rswm_q, hunan$GDPPC))
lag.res <- as.data.frame(lag.list)
colnames(lag.res) <- c("NAME_3", "lag GDPPC")
hunan <- left_join(hunan,lag.res)
head(hunan)
```

```{r}
gdppc <- qtm(hunan, "GDPPC")
lag_gdppc <- qtm(hunan, "lag GDPPC")
tmap_arrange(gdppc, lag_gdppc, asp=1, ncol=2)
```

### **Spatial lag as a sum of neighboring values**

We can calculate spatial lag as a sum of neighboring values by assigning binary weights. This requires us to go back to our neighbors list, then apply a function that will assign binary weights, then we use glist = in the nb2listw function to explicitly assign these weights.

We start by applying a function that will assign a value of 1 per each neighbor. This is done with lapply, which we have been using to manipulate the neighbors structure throughout the past notebooks. Basically it applies a function across each value in the neighbors structure.

```{r}
b_weights <- lapply(wm_q, function(x) 0*x + 1)
b_weights2 <- nb2listw(wm_q, 
                       glist = b_weights, 
                       style = "B")
b_weights2
```

With the proper weights assigned, we can use lag.listw to compute a lag variable from our weight and GDPPC.

```{r}
lag_sum <- list(hunan$NAME_3, lag.listw(b_weights2, hunan$GDPPC))
lag.res <- as.data.frame(lag_sum)
colnames(lag.res) <- c("NAME_3", "lag_sum GDPPC")
```

First, let us examine the result by using the code chunk below.

```{r}
lag_sum
```

Next, we will append the *lag_sum GDPPC* field into `hunan` sf data frame by using the code chunk below.

Now, We can plot both the *GDPPC* and *Spatial Lag Sum GDPPC* for comparison using the code chunk below.

```{r}
hunan <- left_join(hunan, lag.res)
gdppc <- qtm(hunan, "GDPPC")
lag_sum_gdppc <- qtm(hunan, "lag_sum GDPPC")
tmap_arrange(gdppc, lag_sum_gdppc, asp=1, ncol=2)
```

### **Spatial window average**

The spatial window average uses row-standardized weights and includes the diagonal element. To do this in R, we need to go back to the neighbors structure and add the diagonal element before assigning weights.

To add the diagonal element to the neighbour list, we just need to use *include.self()* from **spdep**.

```{r}
wm_qs <- include.self(wm_q)
```

Notice that the Number of nonzero links, Percentage nonzero weights and Average number of links are 536, 6.921488 and 6.090909 respectively as compared to wm_q of 448, 5.785124 and 5.090909

Let us take a good look at the neighbour list of area \[1\] by using the code chunk below.

```{r}
wm_qs[[1]]
```

Notice that now \[1\] has six neighbours instead of five.

Now we obtain weights with *nb2listw()*

```{r}
wm_qs <- nb2listw(wm_qs)
wm_qs
```

Again, we use *nb2listw()* and *glist()* to explicitly assign weight values.

Lastly, we just need to create the lag variable from our weight structure and GDPPC variable.

```{r}
lag_w_avg_gpdpc <- lag.listw(wm_qs, 
                             hunan$GDPPC)
lag_w_avg_gpdpc
```

Next, we will convert the lag variable listw object into a data.frame by using *as.data.frame()*.

```{r}
lag.list.wm_qs <- list(hunan$NAME_3, lag.listw(wm_qs, hunan$GDPPC))
lag_wm_qs.res <- as.data.frame(lag.list.wm_qs)
colnames(lag_wm_qs.res) <- c("NAME_3", "lag_window_avg GDPPC")
```

Note: The third command line on the code chunk above renames the field names of *lag_wm_q1.res* object into *NAME_3* and *lag_window_avg GDPPC* respectively.

Next, the code chunk below will be used to append *lag_window_avg GDPPC* values onto *hunan* sf data.frame by using *left_join()* of **dplyr** package.

```{r}
hunan <- left_join(hunan, lag_wm_qs.res)
```

To compare the values of lag GDPPC and Spatial window average, `kable()` of Knitr package is used to prepare a table using the code chunk below.

```{r}
hunan %>%
  dplyr::select("County", 
         "lag GDPPC", 
         "lag_window_avg GDPPC") %>%
  kable()
```

Lastly, *qtm()* of **tmap** package is used to plot the lag_gdppc and w_ave_gdppc maps next to each other for quick comparison.

```{r}
w_avg_gdppc <- qtm(hunan, "lag_window_avg GDPPC")
tmap_arrange(lag_gdppc, w_avg_gdppc, asp=1, ncol=2)
```

### **Spatial window sum**

The spatial window sum is the counter part of the window average, but without using row-standardized weights.

To add the diagonal element to the neighbour list, we just need to use *include.self()* from **spdep**.

```{r}
wm_qs <- include.self(wm_q)
wm_qs
```

Next, we will assign binary weights to the neighbour structure that includes the diagonal element.

```{r}
b_weights <- lapply(wm_qs, function(x) 0*x + 1)
b_weights[1]
```

Notice that now \[1\] has six neighbours instead of five.

Again, we use *nb2listw()* and *glist()* to explicitly assign weight values.

```{r}
b_weights2 <- nb2listw(wm_qs, 
                       glist = b_weights, 
                       style = "B")
b_weights2
```

With our new weight structure, we can compute the lag variable with *lag.listw()*.

```{r}
w_sum_gdppc <- list(hunan$NAME_3, lag.listw(b_weights2, hunan$GDPPC))
w_sum_gdppc
```

```{r}
w_sum_gdppc.res <- as.data.frame(w_sum_gdppc)
colnames(w_sum_gdppc.res) <- c("NAME_3", "w_sum GDPPC")
hunan <- left_join(hunan, w_sum_gdppc.res)
hunan %>%
  dplyr::select("County", "lag_sum GDPPC", "w_sum GDPPC") %>%
  kable()
w_sum_gdppc <- qtm(hunan, "w_sum GDPPC")
tmap_arrange(lag_sum_gdppc, w_sum_gdppc, asp=1, ncol=2)
```

End.
