---
title: "Take-home Exercise 03"
author: "Leow Xian Zu"
date: "28 Oct 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true 
  message: false #no more warning message
  freeze: true #whatever document already commited does not render
---

This take-home exercise examines the geography of financial inclusion through geographically weighted regression (GWR) to identify and analyse factors influencing access to financial services. Using the FinScope 2023 dataset for Tanzania, this study will focus on district-level insights, offering a spatial perspective on financial inclusion determinants. The exercise will involve geospatial data wrangling, model diagnostics, and geovisualisation, adhering to the grading criteria for data handling, analytical rigour, effective visual communication, and reproducibility in a Quarto environment. The research is grounded in existing literature on financial inclusion, specifically insights from Tanzania, where financial inclusion plays a pivotal role in economic empowerment and reducing income inequality. The goal is to provide a clear, data-driven understanding of spatial accessibility to financial services, with the results aimed at informing policies for broader economic inclusivity.

The research paper on financial inclusion in Tanzania offers valuable insights into the determinants, barriers, and impacts of financial inclusion, highlighting the role of mobile banking and formal financial services in improving economic well-being. It identifies education and income as key determinants and recognises geographic constraints---such as distance to financial institutions---as significant barriers to access. This aligns closely with the objectives of the take-home exercise, which seeks to model the spatial aspects of financial inclusion at the district level. By employing geographically weighted regression (GWR), this exercise will expand upon the research findings by focusing specifically on spatial variability in financial access. The geographic emphasis in this study offers a nuanced understanding of how location influences financial inclusion, which could lead to targeted interventions to overcome geographic barriers and support underserved regions, thereby complementing the study's broader socio-economic conclusions.

# Understanding Tanzania: A Contextual Introduction

As a Singaporean student analysing financial inclusion in Tanzania, I find it crucial to first understand the country's unique characteristics. While Singapore and Tanzania might seem vastly different, both share a British colonial history and gained independence in the 1960s. However, their development paths have diverged significantly. Here's a comprehensive overview of Tanzania that will help frame my analysis:

## Why This Context Matters

Before diving into financial inclusion statistics and analysis, understanding Tanzania's geography, economy, and demographics is essential because:

1\. Physical geography influences access to services

2\. Economic activities affect financial needs

3\. Population distribution impacts service delivery

4\. Infrastructure development determines financial service reach

Understanding these features also helps me:

1\. Identify potential barriers to financial inclusion

2\. Understand regional variations in service access

3\. Appreciate the role of mobile money in overcoming geographical challenges

4\. Recognise why different regions might need different financial solutions

Coming from Singapore's context of universal banking access and high technological adoption, this understanding helps me approach the analysis with appropriate context and avoid making assumptions based on my Singaporean experience. The vast differences in geography, population distribution, and economic activities between Tanzania and Singapore highlight why financial inclusion solutions that work in Singapore might not be directly applicable to Tanzania.

## Key Comparisons with Singapore

```{r}
library(knitr)
library(kableExtra)

comparison_df <- data.frame(
  Aspect = c("Land Area", "Population", "GDP per capita", 
             "Urbanisation", "Main Economic Sectors", "Capital City"),
  Tanzania = c("945,087 km²", "~61 million", "~$1,300 USD",
               "35.2% urban", "Agriculture, Mining, Tourism", 
               "Dodoma (official), Dar es Salaam (de facto)"),
  Singapore = c("728 km²", "~5.7 million", "~$75,000 USD",
                "100% urban", "Finance, Technology, Trade", "Singapore")
)

kable(comparison_df, 
      caption = "Key Comparisons between Tanzania and Singapore") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE) %>%
  column_spec(1, bold = TRUE) %>%
  column_spec(2:3, width = "20em")
```

## Geographical Diversity

Unlike Singapore's uniform urban landscape, Tanzania presents complex geographical features:

1\. Physical Features:

-   800km Indian Ocean coastline

-   Great Rift Valley running through central regions

-   Mount Kilimanjaro (Africa's highest peak)

-   Major lakes (Victoria, Tanganyika)

2\. Economic Zones:

-   Northern Circuit (Tourism)

-   Southern Highlands (Agriculture)

-   Lake Zone (Fishing, Mining)

-   Coastal Zone (Trade, Services)

## Development Challenges

As a student from Singapore, I notice several contrasts in development challenges:

1\. Infrastructure:

-   Transportation networks concentrated in certain regions

-   Rural-urban connectivity issues

-   Varying quality of telecommunications coverage

2\. Economic:

-   Large rural population (64.8%)

-   Regional economic disparities

-   Heavy reliance on agriculture

-   Informal sector significance

3\. Financial Services:

-   45 licensed banks (mostly in urban areas)

-   32.3 million mobile money accounts

-   65% financial inclusion rate

-   Over 100 microfinance institutions

## Administrative Structure

Tanzania's governance structure affects service delivery:

\- 31 regions

\- 184 districts

\- Two capital cities:

\* Dodoma (Official capital, centrally located)

\* Dar es Salaam (Economic hub, coastal location)

In my subsequent analysis, I'll refer back to these contextual factors to ensure my interpretation of financial inclusion patterns is grounded in Tanzania's unique circumstances rather than Singapore's standards.

# Load the packages and data

```{r}
# Load the required packages
pacman::p_load(sf, tidyverse, tmap)

# Read the shapefile
# Replace the path with your actual file location
tz_boundaries <- st_read(dsn = "data/geospatial/", 
                        layer = "geoBoundaries-TZA-ADM2")

# Check the CRS (Coordinate Reference System)
st_crs(tz_boundaries)

# Take a quick look at the data
glimpse(tz_boundaries)

# Create a quick plot to verify the import
tm_shape(tz_boundaries) +
  tm_borders()
```

```{r}
# Load required packages if not already loaded
pacman::p_load(sf, tidyverse, tmap)

# Read the shapefile and explore its contents
tz_boundaries <- st_read("data/geospatial/geoBoundaries-TZA-ADM2.shp")

# Examine the structure and attributes
glimpse(tz_boundaries)

# View column names
names(tz_boundaries)

# Get a summary of the data
summary(tz_boundaries)

# Create an initial visualization with district names
tm_shape(tz_boundaries) +
  tm_borders() +
  tm_layout(main.title = "Tanzania Districts",
            main.title.size = 1)

# Check the number of districts
cat("Number of districts:", nrow(tz_boundaries))
```

```{r}
# Clean and transform the spatial data
tz_districts <- tz_boundaries %>%
  # Keep only necessary columns
  select(district_name = shapeName,
         district_id = shapeID,
         geometry) %>%
  # Convert to more appropriate projection for Tanzania
  st_transform(crs = 32737) %>%  # UTM Zone 37S
  # Arrange alphabetically by district name
  arrange(district_name)

# Create a more detailed map to verify the data
tm_shape(tz_districts) +
  tm_polygons(col = "whitesmoke",
             border.col = "gray30",
             border.alpha = 0.5) +
  tm_layout(main.title = "Tanzania Districts",
            main.title.size = 1,
            frame = FALSE) +
  tm_compass(position = c("right", "top")) +
  tm_scale_bar(position = c("left", "bottom"))

# Save the cleaned spatial data
st_write(tz_districts, 
         "data/cleaned/tz_districts.shp", 
         append = FALSE)

# Create a data frame of district names for reference
district_names <- tz_districts %>%
  st_drop_geometry() %>%
  select(district_name) %>%
  arrange(district_name)

# View first few district names
head(district_names, 10)
```

Some important notes:

1.   I've transformed the CRS to UTM Zone 37S (EPSG:32737) which is more appropriate for Tanzania as it:

    -   Preserves area measurements

    -   Provides more accurate distance calculations

    -   Is suitable for mapping at district level
